---
title: "Hotel Bookings"
author: "illy"
date: "11/26/2021"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

# Prepare
## Import libraries
```{r}
library(tidyverse)
library(skimr)
library(janitor)
library(lubridate)
library(dplyr)
getwd()
```

# Process
## Load the data set
```{r}
hotels <- read.csv("hotel_bookings.csv")
```

## Inspect the data
```{r}
glimpse(hotels)
colnames(hotels)
str(hotels)
```

## Conduct summary ststistics
```{r}
hotels %>%
  summarise(average_lead_time = mean(lead_time),
            max_lead_time = max(lead_time),
            min_lead_time = min(lead_time))
```

# Analyze 
## Total bookings per hotel type
```{r}
hotels %>%
  group_by(hotel, is_canceled) %>%
  filter(is_canceled == 0) %>%
  summarise(bookings = n(), .groups = 'drop')
```
## Order the months
```{r}
hotels$arrival_date_month <- ordered(hotels$arrival_date_month, 
                                     levels = c('January','February','March','April',
                                                'May','June','July','August','September',
        'October','November','December'))
```
## Order the days
```{r}
hotels$arrival_date_day_of_month <- ordered(hotels$arrival_date_day_of_month, 
                                            levels = c('Monday','Tuesday','Wednesday',
                                                       'Thursday','Friday','Saturday',
                                                       'Sunday'))
```

## Number of bookings by month
```{r}
hotels %>%
  group_by(is_canceled, arrival_date_month) %>%
  filter(is_canceled == 0) %>%
  summarise(num_of_bookings = n(), .groups = 'drop') %>%
  arrange(arrival_date_month)
```

## Number of bookings by day
```{r}
hotels %>%
  group_by(is_canceled, arrival_date_day_of_month) %>%
  filter(is_canceled == 0) %>%
  summarise(bookings_per_day = n(), .groups = 'drop') %>%
  arrange(arrival_date_day_of_month)
```

## Lead time by hotel type
```{r}
hotels %>%
  group_by(hotel)%>%
  summarise(average_lead_time = mean(lead_time), .groups = 'drop')
```

## Hotel type booking by month
```{r}
hotels %>%
  filter(is_canceled == 0) %>%
  group_by(hotel, is_canceled, arrival_date_month) %>%
  summarise(booking_per_hotel = n(), .groups = 'drop') %>%
  arrange(arrival_date_month)
```

## Filter the hotel types into different data frames
```{r}
city_hotel <- filter(hotels, hotels$hotel == "City Hotel")
resort <- filter(hotels, hotels$hotel == "Resort Hotel")
```

# Further analysis with visualization
## People with children check in lead time
```{r}
ggplot(data=hotels)+
  geom_point(mapping = aes(x=lead_time,y=children))

```

## People with children weekend stay
```{r}
ggplot(data=hotels)+
  geom_point(mapping = aes(x=stays_in_weekend_nights, y=children))
```

## Distribution channel
```{r}
ggplot(data = hotels) +
  geom_bar(mapping = aes(x = distribution_channel))
```

## Market segment distribution channel
```{r}
ggplot(data = hotels) +
  geom_bar(mapping = aes(x = distribution_channel, fill=market_segment))
```

## Distribution channel by deposit type
```{r}
ggplot(data = hotels) +
  geom_bar(mapping = aes(x = distribution_channel)) +
  facet_grid(~deposit_type) +
  theme(axis.text.x = element_text(angle = 45))
```

## Market segment by hotel type
```{r}
ggplot(data = hotels) +
  geom_bar(mapping = aes(x = hotel, fill = market_segment))
```

## Market segment
```{r}
ggplot(data = hotels) +
  geom_bar(mapping = aes(x = hotel)) +
  facet_wrap(~market_segment)
```

## Online TA market segment 
```{r}
onlineta_city_hotels_v2 <- hotels %>%
  filter(hotel=="City Hotel") %>%
  filter(market_segment=="Online TA")
```

## Minimum and maximum dates 
```{r}
mindate <- min(hotels$arrival_date_year)
maxdate <- max(hotels$arrival_date_year)
```

## Market segment comparison by hotel 
```{r}
ggplot(data = hotels) +
  geom_bar(mapping = aes(x = market_segment)) +
  facet_wrap(~hotel) +
  theme(axis.text.x = element_text(angle = 45)) +
  labs(title="Comparison of market segments by hotel type for hotel bookings",
       caption=paste0("Data from: ", mindate, " to ", maxdate),
       x="Market Segment",
       y="Number of Bookings")
```

